Canonical rings of $\bq$-divisors on minimal rational surfaces

### 2016-07-21 surfaces-write-up/MRL/MRL-surfaces-responses-to-referee.tex

\def\listtodoname{List of Todos}
\def\listoftodos{\@starttoc{tdo}\listtodoname}

\title{Summary of Changes in response to Referee's Report}
\author{Aaron Landesman, Peter Ruhm, and Robin Zhang}

\begin{document}

\maketitle
\section{Changes Made}
Here are the changes we have made in response to the referee's suggestions.
\todo{make sure the theorem number here reflect the new numberings}
\begin{enumerate}
	\item \todo{Ask David how the new version sounds} We explained
		in more detail how our potential application
		for Hilbert modular forms and the difficulty in producing
		an explicit example.
	\item In all of the main theorems, it says that $\frac{c_i}{k_i}$ is in reduced form. 
	\item In Theorem \autoref{thm:proj-generators-relations}, $k$ is a field. Throughout our paper, the results hold for arbitrary fields, but we may assume the field is algebraically closed because the generator and relation degrees are preserved by base change to the algebraic closure. We have now included and clarified this in the ``Notations and Conventions'' subsection at the end of the Introduction.
	\item On page 2, line -9 of the previous version, $\deg(D)$ should have been $\frac{1}{2 \prod_{i=0}^n q_i}$ and is now corrected.
	\item Theorem 1.4 (now \autoref{thm:hirz-generators-relations}) has been rewritten into a simpler form by reducing some
	notation unnecessary to understanding the general idea.
	\item Subsection 1.2 has now been moved to immediately precede Subsection 1.1 as a standalone paragraph.
	\item The reference to later items given on page 5, line 7 of the previous version
		has been moved to Remark \autoref{rem:proj-grobner}.
\item The use of best lower approximations has been replaced by the use of
convergents of Hirzebruch-Jung continued fractions which are equivalent and
well-known.
\item The references for the O'Dorney and Landesman--Ruhm--Zhang papers have been updated. The citation for arxiv articles now includes the arxiv reference numbers.
\item The conditions of Proposition~\autoref{prop:pm-span-and-basis} have been reorganized for clarity.
\item Remark 1.3 has been split into what is now Remark~\autoref{rem:proj-tight-bounds-eff} and Remark~\autoref{rem:proj-tight-bounds} (with
some material moved to Remark~\autoref{rem:exact-noneff-bounds}).

### 2016-07-21 surfaces-write-up/surface-write-up.tex
surfaces: split tightness rem, move notation ssec

\ssec{Main Results and Outline}

@@

@@

\begin{rem}\label{rem:proj-tight-bounds} Note that the bounds
	given in Theorem ~\ref{thm:proj-effective-intro}
are tight and are typically attained, as explained in
Remark ~\ref{rem:exact-effective-bounds}.
Additionally, the bounds given in Theorem 
~\ref{thm:proj-generators-relations} are asymptotically
tight to within a factor of two 
for the following class of divisors. Consider
a divisor 
$ D = \sum_{i=0}^n\frac{p_i}{2q_i}H_i \in \bq \otimes_\bz \di \bp^m$
such that $H_i$ are hyperplanes,  
$q_i$ are pairwise coprime integers, and $p_i$ are chosen
so that $\deg D = \frac{1}{2 \prod_{i=0}^n q_i}$. Further,
choose a linear subspace $\pi: \bp^1 \rightarrow \bp^m$  generically so that
$\pi^* D = \sum_{i=0}^n \frac{p_i}{2q_i}P_i$, 
where $P_i$ are distinct points in $\bp^1$. To choose such a map
$\pi$, we may need to assume that the base field is infinite.
Note that by the Remark immediately following Theorem 8 in 
\cite{dorney:canonical},
the given bounds on the generators and relations of
$R_{\pi^*D}$ are within
a factor of two of the degree of generation and relations of
$R_{\pi^*D}$. Finally, since restriction map
$R_D \rightarrow R_{\pi^*D}$ induced by the restriction maps
on cohomology $H^0(\bp^m, dD) \rightarrow H^0(\bp^1, \pi^*(dD))
\cong H^0(\bp^1, d\pi^* D)$ are surjective, we obtain that
the bounds for the generators and relations of $R_D$ given in
Theorem ~\ref{thm:proj-generators-relations} also
agree with the degree of generation and relations 
to within a factor of two.
 \end{rem}

@@

\begin{restatable}{thm}{hirzrestate}
\label{thm:hirz-generators-relations}
Let $D = \sum_{i = 1}^n \alpha_i D_i \in \bq
\otimes_\bz \di \hirz_m$ where $\alpha_i = \frac{c_i}{k_i} \in \bq$ is written in reduced form.
Let $u, v, z, w$ be the coordinates for the Hirzebruch surface
$\hirz_m$, as described at the beginning of Section
~\ref{sec:hirz} and suppose that $\{f_1, \ldots, f_n\}$
contains two independent 
linear polynomials in $u, v$ and two independent linear polynomials in
$w, x$ {\rm(}with corresponding $\alpha_j$ possibly zero{\rm)}.

@@

\noindent
with relations generated in degrees at most 
\[
	2\rho
\]
\end{restatable}
\begin{rem}
Theorem \ref{thm:hirz-generators-relations-full} provides a more precise bound.
\end{rem}

@@

\ssec{Notations and Conventions}
We will use $m \in \bz$ to index the dimension of a given projective
space $\bp^m$ or the type of the Hirzebruch surface $\hirz_m$.
Let $D$ be a (Weil) $\bq$-divisor of the form
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i \in \bq \otimes_\bz \di X.
\end{align*}\noindent
where $n \in \bz$ indexes the number of irreducible divisors in 
the above expansion of $D$, $\alpha_i \in \bq$, and $D_i \in \di X$ is an integral
codimension 1 closed subscheme of $X$. When it is convenient to do
so, we shall sometimes start the indexing at $0$, so that $i$ runs
from $0$ to $n$. For such a divisor $D$ with $\alpha_i =
\frac{c_i}{k_i}$ let $\ell_i = \lcm_{j\ne i}(k_j)$
and let $\ell_{i,j} = \lcm_{h\ne i,j}(k_h).$ In the case $X = \bp^m$, define the degree of $D$ by $\deg D :=
\sum_{i=1}^{n}\alpha_i \cdot \deg D_i$. The floor of a $\bq$-divisor $D$ is a divisor $\lfloor D
\rfloor \in \di X$ is defined to be $\lfloor D \rfloor := \sum_{i = 1}^{n}
\lfloor \alpha_i \rfloor D_i$. Let $
	R_D := \bigoplus_{d \geq 0} u^d H^0(\bp^m, \lfloor dD \rfloor)$
denote the canonical ring associated to the $\bq$-divisor $D$. If $R$
is a graded ring, we denote the $d$th graded component of $R$ by $R_d$. If $r$ is a rational number, we let
$\fr(r) := r - \lfloor r
\rfloor$ denote the fractional part of $r$. 
Throughout, we work over a fixed algebraically closed field $\bk$.
Since generator and relation degrees are preserved under base change to the algebraic closure, assuming that $\bk$ is algebraically closed is no restriction
on generality and the results hold equally well over an arbitrary field.

If $D \in \bq \otimes_\bz
\di X$ is an arbitrary divisor, we denote $h^0(X, D) := \dim_\bk H^0(X,D)$.

@@

>>>>

We will use $m \in \bz$ to index the dimension of a given projective
space $\bp^m$ or the type of the Hirzebruch surface $\hirz_m$.
Let $D$ be a (Weil) $\bq$-divisor of the form
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i \in \bq \otimes_\bz \di X.
\end{align*}\noindent
where $n \in \bz$ indexes the number of irreducible divisors in 
the above expansion of $D$, $\alpha_i \in \bq$, and $D_i \in \di X$ is an integral
codimension 1 closed subscheme of $X$. When it is convenient to do
so, we shall sometimes start the indexing at $0$, so that $i$ runs
from $0$ to $n$. In the case $X = \bp^m$, define the degree of $D$ by $\deg D :=
\sum_{i=1}^{n}\alpha_i \cdot \deg D_i$. The floor of a $\bq$-divisor $D$ is a divisor $\lfloor D
\rfloor \in \di X$ is defined to be $\lfloor D \rfloor := \sum_{i = 1}^{n}
\lfloor \alpha_i \rfloor D_i$. Let $
	R_D := \bigoplus_{d \geq 0} u^d H^0(\bp^m, \lfloor dD \rfloor)$
denote the canonical ring associated to the $\bq$-divisor $D$. If $R$
is a graded ring, we denote the $d$th graded component of $R$ by $R_d$. If $r$ is a rational number, we let
$\fr(r) := r - \lfloor r
\rfloor$ denote the fractional part of $r$. 
Throughout, we work over a fixed algebraically closed field $\bk$.
Since generator and relation degrees are preserved under base change to the algebraic closure, assuming that $\bk$ is algebraically closed is no restriction
on generality and the results hold equally well over an arbitrary field.
If $D \in \bq \otimes_\bz
\di X$ is an arbitrary divisor, we denote $h^0(X, D) := \dim_\bk H^0(X,D)$.

\ssec{Main Results and Outline}
\label{ssec:main}

@@

\begin{rem}
\label{rem:proj-tight-bounds-eff}
Note that the bounds
given in Theorem ~\ref{thm:proj-effective-intro}
are tight and are typically attained, as explained in
Remark ~\ref{rem:exact-effective-bounds}.
\end{rem}

@@

\begin{rem}
\label{rem:proj-tight-bounds}
Note that the bounds given in Theorem 
~\ref{thm:proj-generators-relations} are asymptotically
tight to within a factor of two 
for a class of divisors described in
Remark~\ref{rem:exact-noneff-bounds}.
\end{rem}

@@

\begin{thm}
\label{thm:hirz-generators-relations}
Let $D = \sum_{i = 1}^n \alpha_i D_i \in \bq
\otimes_\bz \di \hirz_m$ where $\alpha_i = \frac{c_i}{k_i} \in \bq$ is written in reduced form.
Let $u, v, z, w$ be the coordinates for the Hirzebruch surface
$\hirz_m$, as described at the beginning of Section
~\ref{sec:hirz} and suppose that $\{f_1, \ldots, f_n\}$
contains two independent 
linear polynomials in $u, v$ and two independent linear polynomials in
$w, x$ {\rm(}with corresponding $\alpha_j$ possibly zero{\rm)}.

@@

\noindent
with relations generated in degrees at most 
\[
	2\rho
\]
\end{thm}

\begin{rem}
Theorem~\ref{thm:hirz-generators-relations} is restated with a more precise
bound in Theorem~\ref{thm:hirz-generators-relations-full}.
\end{rem}

@@

@@

\begin{rem}
\label{rem:exact-noneff-bounds}
The bounds given in Theorem 
~\ref{thm:proj-generators-relations} are asymptotically
tight to within a factor of two 
for the following class of divisors. Consider
a divisor 
$ D = \sum_{i=0}^n\frac{p_i}{2q_i}H_i \in \bq \otimes_\bz \di \bp^m$
such that $H_i$ are hyperplanes,  
$q_i$ are pairwise coprime integers, and $p_i$ are chosen
so that $\deg D = \frac{1}{2 \prod_{i=0}^n q_i}$. Further,
choose a linear subspace $\pi: \bp^1 \rightarrow \bp^m$  generically so that
$\pi^* D = \sum_{i=0}^n \frac{p_i}{2q_i}P_i$, 
where $P_i$ are distinct points in $\bp^1$. To choose such a map
$\pi$, we may need to assume that the base field is infinite.
Note that by the Remark immediately following Theorem 8 in 
\cite{dorney:canonical},
the given bounds on the generators and relations of
$R_{\pi^*D}$ are within
a factor of two of the degree of generation and relations of
$R_{\pi^*D}$. Finally, since restriction map
$R_D \rightarrow R_{\pi^*D}$ induced by the restriction maps
on cohomology $H^0(\bp^m, dD) \rightarrow H^0(\bp^1, \pi^*(dD))
\cong H^0(\bp^1, d\pi^* D)$ are surjective, we obtain that
the bounds for the generators and relations of $R_D$ given in
Theorem ~\ref{thm:proj-generators-relations} also
agree with the degree of generation and relations 
to within a factor of two.
\end{rem}
